---
title: "Possible QTL identification and bayesian significant intervals continued"
subtitle: 'Report 6'
author: 'Kaleb Negussie' 
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    fig_caption: yes
    fig_retiSB: 1 
    code_folding: show
date: '2022-07-18'
editor_options: 
  markdown: 
    wrap: 72
---

```{r load-packages, include=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(ggthemes)
library(knitr)
library(qtl)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
df <- read.cross("csvr", ".", "gen.csv", na.strings = 'NA')
```

```{r}
ClS.out<-scanone(df, pheno.col = "Cl_Salt",method="mr")
KS.out<-scanone(df, pheno.col = "K_Salt",model = 'normal',method="mr")
NaS.out<-scanone(df, pheno.col = "Na_Salt", method="mr")
SBS.out<-scanone(df, pheno.col = "SB_Salt",model = 'normal', method="mr")
KW.out<-scanone(df, pheno.col = "K_Water",model = 'normal', method="mr")
ClW.out<-scanone(df,pheno.col = "Cl_Water",model = 'normal', method="mr")
NaW.out<-scanone(df,pheno.col = "Na_Water",model = 'normal', method="mr")
SBW.out <- scanone(df,pheno.col = "SB_Water",model = 'normal', method="mr")
```

```{r Permutation Test, message=FALSE, warning=FALSE, eval=FALSE}


set.seed(287349723)
ClW.perm <- scanone(pheno.col = ('Cl_Water'), df, method = 'mr', n.perm = 1000)
ClS.perm <- scanone(pheno.col = ('Cl_Salt'), df, method = 'mr', n.perm = 1000)
KS.perm <- scanone(pheno.col = ('K_Salt'), df, method = 'mr', n.perm = 1000)
KW.perm <- scanone(pheno.col = ('K_Water'), df, method = 'mr', n.perm = 1000)
NaS.perm <- scanone(pheno.col = ('Na_Salt'), df, method = 'mr', n.perm = 1000)
NaW.perm <- scanone(pheno.col = ('Na_Water'), df, method = 'mr', n.perm = 1000)
SBS.perm <- scanone(pheno.col = ('Cl_Water'), df, method = 'mr', n.perm = 1000)
SBW.perm <- scanone(pheno.col = ('Cl_Water'), df, method = 'mr', n.perm = 1000)


saveRDS(ClW.perm, file= "/Users/kalebnegussie/Desktop/RI data/ClW_perm")
saveRDS(ClS.perm, file= "/Users/kalebnegussie/Desktop/RI data/ClS_perm")
saveRDS(KS.perm, file= "/Users/kalebnegussie/Desktop/RI data/KS_perm")
saveRDS(KW.perm, file= "/Users/kalebnegussie/Desktop/RI data/KW_perm")
saveRDS(NaS.perm, file= "/Users/kalebnegussie/Desktop/RI data/NaS_perm")
saveRDS(NaW.perm, file= "/Users/kalebnegussie/Desktop/RI data/NaW_perm")
saveRDS(SBS.perm, file= "/Users/kalebnegussie/Desktop/RI data/SBS_perm")
saveRDS(SBW.perm, file= "/Users/kalebnegussie/Desktop/RI data/SBW_perm")

```

# Chlorine(Cl)

## Water Diet

### Permutation Test

```{r}
ClW.perm <- readRDS("/Users/kalebnegussie/Desktop/RI data/ClW_perm")
threshClW <- summary(ClW.perm, c(alpha = 0.67, alpha = 0.10, alpha = 0.05))
plot(ClW.perm, main = 'Cl(Water)')
abline(v = threshClW[3], lty= "dotted",lwd = 1, col="red")
```

### Mainscan

*Blue line represents 67% suggestive threshold*

```{r}
plot(ClW.out, main = 'Cl(Water)', bandcol="gray85", alternate.chrid= TRUE)
abline(h = threshClW[1], lty= "dotted",lwd = 1, col="blue")
abline(h = threshClW[2], lty= "dotted",lwd = 1, col="red")
abline(h = threshClW[3], lty= "dotted",lwd = 1, col="green")
```

### Phenotype Against Genotype

*2 possible QTLs reached the suggestive threshold*

```{r}
set.seed(234423)
summary(ClW.out, perms=ClW.perm, alpha= 0.67, pvalues=TRUE)

plotPXG(df,marker= 'rat108_035_d11.q1ca_316',ylab = "", main = "Clorine-Water Possible QTL(rat108_035_d11.q1ca_316)" )

plotPXG(df,marker= 'J700466',ylab = "", main = "Clorine-Water Possible QTL(J700466)" )



```

### Confidence Values

```{r}
ClW.bayes <- bayesint(ClW.out, chr = 1)
ClW.bayes2 <- bayesint(ClW.out, chr = 14)



ClW.CV <- matrix(c(1, 0.31, round(ClW.bayes[2,3],digits = 2),
                   round(ClW.bayes[2,2],2), round(ClW.bayes[1,2],2), round(ClW.bayes[3,2],2),
                   14, 0.41, round(ClW.bayes2[2,3],digits = 2),
                   round(ClW.bayes2[2,2],2), round(ClW.bayes2[1,2],2), round(ClW.bayes2[3,2],2)),
                 ncol=6, byrow=TRUE)
colnames(ClW.CV) <- c('chr','P-value','LOD','position(CM)', 'bayesint(lower)','bayesint(upper)')
rownames(ClW.CV) <- c('rat108_035...', 'J700466')
ClW.CV
```

## Salt Diet

### Permutation Test

```{r}
ClS.perm <- readRDS("/Users/kalebnegussie/Desktop/RI data/ClS_perm")
threshClS <- summary(ClS.perm, c(alpha = 0.67, alpha = 0.10, alpha = 0.05))
plot(ClS.perm, main = 'Cl(Salt)')
abline(v = threshClS[3], lty= "dotted",lwd = 1, col="red")
```

### Main-scan

*Blue line represents 67% suggestive threshold*

```{r}
plot(ClS.out, main = 'Cl(Salt)', bandcol="gray85", alternate.chrid= TRUE)
abline(h = threshClS[1], lty= "dotted",lwd = 1, col="blue")
abline(h = threshClS[2], lty= "dotted",lwd = 1, col="red")
abline(h = threshClS[3], lty= "dotted",lwd = 1, col="green")
```

### Phenotype Against Genotype

*One possible QTL reached the suggestive threshold*

```{r}
set.seed(234423)
summary(ClS.out, perms=ClS.perm, alpha= 0.67, pvalues=TRUE)
plotPXG(df,marker= 'J571386',ylab = "", main = "Clorine-Salt Possible QTL(J571386)" )
```

### Confidence Values

```{r}
ClS.bayes <- bayesint(ClS.out, chr = 3)
ClS.CV <- matrix(c(3, 0.60, round(ClS.bayes[2,3],digits = 2),
                   round(ClS.bayes[2,2],2), round(ClS.bayes[1,2],2), round(ClS.bayes[3,2],2)),
                 ncol=6, byrow=TRUE)
colnames(ClS.CV) <- c('chr','P-value','LOD','position(CM)', 'bayesint(lower)','bayesint(upper)')
rownames(ClS.CV) <- c('J571386')
ClS.CV
```

## Diet Comparison

\*Salt diet in red

```{r}
plot(ClS.out, ClW.out, main = 'Cl Salt vs. Water QTL plot',
     col= c('red','blue'), chr = c(1,2,7,8,10,14), bandcol="gray85" )

```

# Potassium(K)

## Water Diet

### Permutation Test

```{r}
KW.perm <- readRDS("/Users/kalebnegussie/Desktop/RI data/KW_perm")
threshKW <- summary(KW.perm, c(alpha = 0.67, alpha = 0.10, alpha = 0.05))
plot(KW.perm, main = 'K(Water)')
abline(v = threshKW[3], lty= "dotted",lwd = 1, col="red")
```

### Main-scan

*Blue line represents 67% suggestive threshold*

```{r}
plot(KW.out, main = 'K(Water)', bandcol="gray85", alternate.chrid= TRUE)
abline(h = threshKW[1], lty= "dotted",lwd = 1, col="blue")
abline(h = threshKW[2], lty= "dotted",lwd = 1, col="red")
abline(h = threshKW[3], lty= "dotted",lwd = 1, col="green")
```

### Phenotype Against Genotype

*There were no possible QTLs that reached the suggestive threshold*

```{r}
set.seed(234423)
summary(KW.out, perms=KW.perm, alpha= 0.67, pvalues=TRUE)
```

## Salt Diet

### Permutation Test

```{r}
KS.perm <- readRDS("/Users/kalebnegussie/Desktop/RI data/KS_perm")
threshKS <- summary(KS.perm, c(alpha = 0.67, alpha = 0.10, alpha = 0.05))
plot(KS.perm, main = 'K(Salt)')
abline(v = threshKS[3], lty= "dotted",lwd = 1, col="red")
```

### Main-scan

*Blue line represents 67% suggestive threshold*

```{r}
plot(KS.out, main = 'K(Salt)', bandcol="gray85", alternate.chrid= TRUE)
abline(h = threshKS[1], lty= "dotted",lwd = 1, col="blue")
abline(h = threshKS[2], lty= "dotted",lwd = 1, col="red")
abline(h = threshKS[3], lty= "dotted",lwd = 1, col="green")
```

### Phenotype Against Genotype

*There were no possible QTLs that reached the suggestive threshold*

```{r}
set.seed(234423)
summary(KS.out, perms=KS.perm, alpha= 0.67, pvalues=TRUE)
```

## Diet Comparison

*Salt diet in red*

```{r}
plot(KS.out, KW.out, main = 'K Salt vs. Water QTL plot', 
     col= c('red','blue'), chr = c(1,2,4,5,6,9), bandcol="gray85")

```

# Sodium(Na)

## Water Diet

### Permutation Test

```{r}
NaW.perm <- readRDS("/Users/kalebnegussie/Desktop/RI data/NaW_perm")
threshNaW <- summary(NaW.perm, c(alpha = 0.67, alpha = 0.10, alpha = 0.05))
plot(NaW.perm, main = 'Na(Water)')
abline(v = threshNaW[3], lty= "dotted",lwd = 1, col="red")
```

### Main-scan

*Blue line represents 67% suggestive threshold*

```{r}
plot(NaW.out, main = 'Na(Water)', bandcol="gray85", alterNate.chrid= TRUE)
abline(h = threshNaW[1], lty= "dotted",lwd = 1, col="blue")
abline(h = threshNaW[2], lty= "dotted",lwd = 1, col="red")
abline(h = threshNaW[3], lty= "dotted",lwd = 1, col="green")
```

### Phenotype Against Genotype

*3 possible QTLs reached the suggestive threshold*

```{r}
summary(NaW.out, perms=NaW.perm, alpha= 0.67, pvalues=TRUE)
plotPXG(df,marker= 'rdahl-10d19_fp2_b1_135',ylab = "", 
        main = "Sodium-Water Possible QTL(rdahl-10d19_fp2_b1_135)" )

plotPXG(df,marker= 'rat104_080_h22.q1ca_403',ylab = "", 
        main = "Sodium-Water Possible QTL(rat104_080_h22.q1ca_403)" )

plotPXG(df,marker= 'rat109_060_e09.q1ca_157',ylab = "", 
        main = "Sodium-Water Possible QTL(rat109_060_e09.q1ca_157)" )


```

### Confidence Values

```{r}
NaW.bayes <- bayesint(NaW.out, chr = 2)
NaW.bayes2 <- bayesint(NaW.out, chr = 6)
NaW.bayes3 <- bayesint(NaW.out, chr = 17)

NaW.CV <- matrix(c(2, 0.35, round(NaW.bayes[2,3],digits = 2),
                   round(NaW.bayes[2,2],2), round(NaW.bayes[1,2],2), round(NaW.bayes[3,2],2),
                   6, 0.49, round(NaW.bayes2[2,3],digits = 2),
                   round(NaW.bayes2[2,2],2), round(NaW.bayes2[1,2],2), round(NaW.bayes2[3,2],2),
                   17, 0.42, round(NaW.bayes3[2,3],digits = 2),
                   round(NaW.bayes3[2,2],2), round(NaW.bayes3[1,2],2), round(NaW.bayes3[3,2],2)),
                 ncol=6, byrow=TRUE)
colnames(NaW.CV) <- c('chr','P-value','LOD','position(CM)', 'bayesint(lower)','bayesint(upper)')
rownames(NaW.CV) <- c('rdahl-10...', 'gko-98h8...', 'rat109...')
NaW.CV
```

## Salt Diet

### Permutation Test

```{r}
NaS.perm <- readRDS("/Users/kalebnegussie/Desktop/RI data/NaS_perm")
threshNaS <- summary(NaS.perm, c(alpha = 0.67, alpha = 0.10, alpha = 0.05))
plot(NaS.perm, main = 'Na(Salt)')
abline(v = threshNaS[3], lty= "dotted",lwd = 1, col="red")
```

### Main-scan

*Blue line represents 67% suggestive threshold*

```{r}
plot(NaS.out, main = 'Na(Salt)', bandcol="gray85", alterNate.chrid= TRUE)
abline(h = threshNaS[1], lty= "dotted",lwd = 1, col="blue")
abline(h = threshNaS[2], lty= "dotted",lwd = 1, col="red")
abline(h = threshNaS[3], lty= "dotted",lwd = 1, col="green")
```

### Phenotype Against Genotype

*One possible QTL reached the suggestive threshold*

```{r}
set.seed(234423)
summary(NaS.out, perms=NaS.perm, alpha= 0.67, pvalues=TRUE)
plotPXG(df,marker= 'rat101_028_i20.p1ca_250',ylab = "", main = "Sodium-Salt Possible QTL(rat101_028_i20.p1ca_250)" )
```

### Confidence Values

```{r}
NaS.bayes <- bayesint(NaS.out, chr = 16)
NaS.CV <- matrix(c(16, 0.60, round(NaS.bayes[2,3],digits = 2),
                   round(NaW.bayes[2,2],2), round(NaS.bayes[1,2],2), round(NaS.bayes[3,2],2)),
                 ncol=6, byrow=TRUE)
colnames(NaS.CV) <- c('chr','P-value','LOD','position(CM)', 'bayesint(lower)','bayesint(upper)')
rownames(NaS.CV) <- c('rat102_02...')
NaS.CV
```

## Diet Comparison

\*Salt diet in red

```{r}
plot(NaS.out,NaW.out, main = 'Na Salt vs. Water QTL plot',
     col= c('red','blue'), chr = c(2,3,6,11,16),bandcol="gray85")


```

# Skin:Body Ratio

## Water Diet

### Permutation Test

```{r}
SBW.perm <- readRDS("/Users/kalebnegussie/Desktop/RI data/SBW_perm")
threshSBW <- summary(SBW.perm, c(alpha = 0.67, alpha = 0.10, alpha = 0.05))
plot(SBW.perm, main = 'SB(Water)')
abline(v = threshSBW[3], lty= "dotted",lwd = 1, col="red")
```

### Main-scan

*Blue line represents 67% suggestive threshold*

```{r}
plot(SBW.out, main = 'SB(Water)', bandcol="gray85", alternate.chrid= TRUE)
abline(h = threshSBW[1], lty= "dotted",lwd = 1, col="blue")
abline(h = threshSBW[2], lty= "dotted",lwd = 1, col="red")
abline(h = threshSBW[3], lty= "dotted",lwd = 1, col="green")
```

### Phenotype Against Genotype

*One possible QTL reached the suggestive threshold*

```{r}
set.seed(234423)
summary(SBW.out, perms=SBW.perm, alpha= 0.67, pvalues=TRUE)
plotPXG(df,marker= 'rat102_024_l24.q1ca_481',ylab = "", 
        main = "Skin:Body-Water Possible QTL(rat102_024_l24.q1ca_481)" )

```

### Confidence Values

```{r}
SBW.bayes <- bayesint(SBW.out, chr = 16)
SBW.CV <- matrix(c(16, 0.57, round(SBW.bayes[2,3],digits = 2),
                   round(SBW.bayes[2,2],2), round(SBW.bayes[1,2],2), round(SBW.bayes[3,2],2)),
                 ncol=6, byrow=TRUE)
colnames(SBW.CV) <- c('chr','P-value','LOD','position(CM)', 'bayesint(lower)','bayesint(upper)')
rownames(SBW.CV) <- c('rat102_02...')
SBW.CV
```

## Salt Diet

### Permutation Test

```{r}
SBS.perm <- readRDS("/Users/kalebnegussie/Desktop/RI data/SBS_perm")
threshSBS <- summary(SBS.perm, c(alpha = 0.67, alpha = 0.10, alpha = 0.05))
plot(SBS.perm, main = 'SB(Salt)')
abline(v = threshSBS[3], lty= "dotted",lwd = 1, col="red")
```

### Main-scan

*Blue line represents 67% suggestive threshold*

```{r}
plot(SBS.out, main = 'SB(Salt)', bandcol="gray85", alternate.chrid= TRUE)
abline(h = threshSBS[1], lty= "dotted",lwd = 1, col="blue")
abline(h = threshSBS[2], lty= "dotted",lwd = 1, col="red")
abline(h = threshSBS[3], lty= "dotted",lwd = 1, col="green")
```

### Phenotype Against Genotype

*No possible QTL reached the suggestive threshold*

```{r}
set.seed(234423)
summary(SBS.out, perms=SBS.perm, alpha= 0.67, pvalues=TRUE)
```

## Diet Comparison

*Salt diet in red*

```{r}
plot(SBS.out,SBW.out, main = 'Skin:Body Salt vs. Water QTL plot', 
     col= c('red','blue'), chr = c(4,5,7,8,10,14,17))

```
