---
title: "RNA-seq data analysis"
subtitle: 'PCA, hierarchical clustering, and dispersion analysis'
author: 'Kaleb Negussie' 
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    fig_caption: yes
    fig_retiSB: 1 
    code_folding: show
date: '2022-11-19'
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, message=FALSE, warning=FALSE}
#Import count table and replaces first row to rownames
library('DESeq2')
count <- read.csv("/Users/kalebnegussie/Desktop/RI data/count_table")
count[,1] <- NULL
rownames(count) <- count[,1]
count[,1] <- NULL
#Creates a filtered data frame where only rna transcripts with more than 50 
#samples across all mice was detected
countfil <- count[rowSums(count[])>50,]
```

```{r setup, message=FALSE, warning=FALSE}
#Builds a treatment table
treatment_table<- matrix(,length(colnames(countfil)),)

Salt <- ".1"

TT <- list()
for (i in colnames(countfil)) {
  for (j in treatment_table[1,]){

  if (grepl(Salt, i, fixed = TRUE)){
  j <- "Salt" }
  else {j <- "Water"}
  }
    TT<- c(TT,j)
  }
  
JJ <- list()
for (i in colnames(countfil)) {
  for (j in treatment_table[1,]){
    i = substr(i,1,nchar(i)-2)
    
    JJ <- c(JJ,i)

}}


treatment_table <- matrix(TT)
treatment_table <- as.matrix.data.frame(treatment_table)

rownames(treatment_table) <- colnames(countfil)
colnames(treatment_table) <- c('Diet')
```

```{r setup, message=FALSE, warning=FALSE}
#Creates unfiltered deSeq object 
#(deSeq automatically removes samples with 0 counts across all columns)

dds_unfil <- DESeqDataSetFromMatrix(countData = round(count),
                             colData = treatment_table,
                             design = ~ Diet)
dds_unfil <- DESeq(dds_unfil)

#Creates filtered dds object
dds_fil <- DESeqDataSetFromMatrix(countData = round(countfil),
                             colData = treatment_table,
                             design = ~ Diet)
dds_fil <- DESeq(dds_fil)

fil_norm <- vst(dds_fil)
unfil_norm <- vst(dds_unfil)
```

# Dispersion Plot Filtered vs. Unfiltered 

There are few values with very high dispersion which are not shrunk
towards The fitted curve. However there were a fair bit of values with
very low dispersion values which were fitted strongly.

```{r message=FALSE, warning=FALSE}
plotDispEsts(dds_fil)
plotDispEsts(dds_unfil)
```

# Principle Component Analysis

The original data did not have multiple batches analyzed for each
strain, therefore batch corrections do not need to be made. No obvious
grouping exists within the treatment group, grouping by strain would
likely show clustering as observed in the dendrogram shown later.

```{r message=FALSE, warning=FALSE}
plotPCA(fil_norm)
```

# Dendrogram

The dendrogram shows strong clustering by strain with the closest
expression assosiations being those of individuals with different diets
with the same strain.

```{r message=FALSE, warning=FALSE}
library(magrittr)
library(RColorBrewer)
rv <- rowVars(assay(dds_vst))
o <- order(rv,decreasing=TRUE)
dists <- dist(t(assay(dds_vst)[head(o,500),]))
hc <- hclust(dists)
plot(hc, labels=dds_vst$sample)

plot(hclust(as.dist(1-cor(countfil, method="spearman"))))

```

#Determining Parameters for WGCNA 
The WGCNA tutorial recommends choosing a power value with an associated r^2 
value above 0.80 to minimize noise. However, all of the power values shown are 
well below the 0.80 threshold. Moving forward a power value of 6 will fairly 
arbitrarily be used.  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(WGCNA)
library(dplyr)
library(ggplot2)
fil_norm <- vst(dds_fil)
normalized_counts <- assay(fil_norm)

sft <- pickSoftThreshold(normalized_counts,
  dataIsExpr = TRUE,
  corFnc = cor,
  networkType = "signed"
)

sft_df <- data.frame(sft$fitIndices) %>%
  dplyr::mutate(model_fit = sign(slope) * SFT.R.sq)

ggplot(sft_df, aes(x = Power, y = model_fit, label = Power)) +
  geom_point() +
  geom_text(nudge_y = 0.1) +
  geom_hline(yintercept = 0.80, col = "red") +
  ylim(c(min(sft_df$model_fit), 1.05)) +
  xlab("Soft Threshold (power)") +
  ylab("Scale Free Topology Model Fit, signed R^2") +
  ggtitle("Scale independence") +
  theme_classic()

```
#WGCNA 

```{r message=FALSE, warning=FALSE}
#Creating Block-wise Network 
#bwnet <- blockwiseModules(normalized_counts,
#  maxBlockSize = 5000, 
#  TOMType = "signed", 
#  power = 6, 
#  numericLabels = TRUE, 
#  randomSeed = 2259,
#)

#Saving block-wise modules desktop
#readr::write_rds(bwnet,
#  file = "/Users/kalebnegussie/Desktop/RI data/bwnet.RDS"
#)

lv<- readRDS("/Users/kalebnegussie/Desktop/RI data/bwnet.RDS")
```

#Count Outlier Detection


The below plot shows Cook's distance over the rank of a test statistic.
Cook's distance is the degree to which the fitted coeffiencts would
change with the removal of a value.

